Proposal: Loose Grammar Inference for Lossy Compression

### 2012-10-18 proposal.tex
edits, and a few more transition sentences

Many sources of data exhibit natural hierarchical structure, but it is
oftentimes difficult to apprehend this structure at a high level due to small
fluctuations in the data at the lower levels of the hierarchy.  If one is
interested primarily in the structure of the data, a lossy compression scheme
has the potential to mitigate some of this low-level noise in favor of the
high-level signal, and can reduce the data size to boot.

@@

We are interested in whether lossy compression techniques can be combined with
CFG transformation and compression techniques like \Sequitur's in order (a) to
illuminate the high-level grammatical structure of the data, and (b) to more
effectively compress the data with some tolerable degree of lossiness.  That
is, our goals are twofold---we wish to develop a technique which can both
increase human understanding of a complex document and yield good compression
ratios.  Notably, if we are interested in the structure rather than the
details, we surmise that the lossiness we introduce can be highly desirable.
For instance, introducing lossiness to \Sequitur's transformation on musical
passages could result in interesting, even music-theoretically sound
reductions, of a given piece.

@@

There seems to be little existing work done in the intersection of grammar
compression and lossy compression.  Thus, we will compare our work to existing
techniques both in lossless hierarchical compression and lossy non-hierarchical
compression.  Moreover, we will not concern ourselves with fitting our lossy
grammar compression into overarching information-theoretic developments.

@@

Though we have not found very much previous work in the area of lossy grammar
based compression, our proposed research will try to build upon powerful
results in grammar based compression and incorporate some of ideas currently
used in lossy compression algorithms.

@@

Starting with an existing grammar inference algorithm, or perhaps with a simple
variation based on the reduction rules in \cite{grammarcodes}, there are three
places we might introduce lossiness into the system.  We could
\begin{itemize}
  \item perform text-transforms on the input before applying a grammar
    inference algorithms;
  \item modify the grammar inference algorithm so that it generates a lossy
    grammar on the fly; or
  \item take a lossless grammar code and apply lossy transforms to it.
\end{itemize}
Of these, the first could provide the biggest potential for taking advantage of
source-specific knowledge of what kinds of data we can ignore.  The second
allows the most control over precisely how a grammar can be formed, and would
not be too difficult, we expect, to build into reduction-rule based frameworks.
The third would be the most general, since it could be applied to any CFG, but
may be difficult to get consistent performance from.

@@

Comparing our runtime with existing algorithms can be done in a straightforward
way.

We plan on evaluating our algorithm on standard english prose texts, but we
also suspect that the texts which will prove most interesting to analyze are
those with slightly more structure imposed on them. Human-made tables such as
the genealogical trees discussed in \cite{nevillphd} are a good example, as are
simple musical pieces. Slightly more artificial examples we might try include
strings generated from actual CFGs but with small amounts of noise added in.

>>>>

Many sources of data exhibit natural hierarchical structure, but it is
often difficult to apprehend this structure at a high level due to small
fluctuations at the lower levels of the hierarchy.  If one is
interested primarily in the structure of the data, a lossy compression scheme
has the potential to mitigate some of this low-level noise in favor of the
high-level signal, and can greatly reduce the compressed data size.

@@

We are interested in whether lossy compression techniques can be combined with
CFG transformation and compression techniques like \Sequitur's in order (a) to
illuminate the high-level grammatical structure of the data, and (b) to more
effectively compress the data with some tolerable degree of lossiness.  That
is, our goals are twofold---we wish to develop a technique which can both
increase human understanding of a complex document and yield good compression
ratios.  Notably, if we are interested in the structure rather than the
details, we surmise that the lossiness we introduce can be highly desirable.
For instance, introducing lossiness to \Sequitur's transformation on musical
passages could result in interesting, even music-theoretically sound, 
reductions of a given piece.

@@

There seems to be little existing work done in the intersection of grammar
compression and lossy compression.  Thus, we will compare our work to existing
techniques both in lossless hierarchical compression and lossy non-hierarchical
compression.  Moreover, we are not too concerned with fitting our lossy
grammar compression into information theoretical developments, since
our main concerns are achieving interesting human comprehensible compressions
with reasonable fidelity and good ratios as a benchmark.

@@

Though we have not found very much previous work in the area of lossy grammar
based compression, our proposed research will try to build upon established
results in grammar based compression and incorporate some of ideas currently
used in lossy compression algorithms.

@@

Starting with an existing grammar inference algorithm, or perhaps with a simple
variation based on the reduction rules in \cite{grammarcodes}, there are three
places we might introduce lossiness into the system.  We could
\begin{itemize}
  \item perform text-transforms on the input before applying grammar
    inference algorithms;
  \item modify the grammar inference algorithm so that it generates a lossy
    grammar on the fly; or
  \item take a lossless grammar code and apply lossy transforms to it.
\end{itemize}
Of these, the first could provide the biggest potential for taking advantage of
source-specific knowledge of what kinds of data we can ignore.  The second
allows the most control over precisely how a grammar can be formed, and would
not be too difficult, we expect, to build into reduction-rule based frameworks.
The third would be the most general, since it could be applied to any CFG, but
may be difficult to get consistent performance from.

@@

Comparing our runtime with existing algorithms can be done in a straightforward
way. Though our main focus will be on smaller datasets with human-readable
structures, we will also experiment to see the effects of lossiness on
scalability.

We plan on evaluating our algorithm on standard english prose texts, but we
also suspect that the texts which will prove most interesting to analyze are
those with slightly more structure imposed on them. Human-made tables such as
the genealogical trees discussed in \cite{nevillphd} are a good example, as are
simple musical pieces. Slightly more artificial examples we might try include
strings generated from actual CFGs but with small amounts of noise added in.
These will reveal best case behavior in compressing a nearly-perfectly
structured dataset.
